---
title: "Reproducible Research Peer Assignment 1"
author: "thenuv"
date: "15 July 2018"
output: html_document 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## "Analysis with Personal movement data"

This analyis is from the personal movement data collected with Activity monitoring devices.


### Download Activity monitoring data
```{r DownloadData }
fileurl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
filename <- "repdataFdataFactivity.zip"
if (!file.exists(filename)) {
        download.file(fileurl, filename)
}
if (file.exists(filename) & !file.exists("activity.csv")){
        unzip(filename)
}
```

### 1.Read Activity data from file
```{r Read File}
df <- read.csv("activity.csv", sep = ",", header= TRUE)
str(df)

```

### 2.Histogram for Total number of steps taken per day
``` {r Histogram }
df$date <- as.Date(df$date)
p2 <- with(df, aggregate(steps ~ date, FUN = sum, na.rm = TRUE))
hist(p2$steps, main="Histogram of Steps Per Day", xlab = "Total number of Steps per day", col = "steelblue")

```


### 3.Display Mean & Median for steps per day
```{r Mean Median}
p3 <- with(df, aggregate(steps ~ date, FUN = mean, na.rm = TRUE))
names(p3) <- c("Date", "Mean")
p3_tmp <- with(df, aggregate(steps ~ date, FUN = median, na.rm = TRUE))
names(p3_tmp) <- c("Date", "Median")
p3 <- merge(p3, p3_tmp, by = "Date")
print(p3, type="HTML")

```


### 4.Time Series Plot for Steps at 5 mintues interval (Averaged per day)
```{r Timeseries Plot}
p4 <- with(df, aggregate(steps ~ interval, FUN = mean, na.rm = TRUE))
with (p4, plot(interval, steps, type="l", col="brown"))
title (main="Average Daily Activity", xlab = "Interval (5 Minutes)", ylab = "Steps")
```

### 5.Interval having maximum number of Steps
```{r MaxSteps }
p5 <- p4[p4$steps==max(p4$steps),]
print(p5)
```

Maximum number of steps **`r p5$steps`** (averaged across all days) was observed at **`r format(strptime(paste("0", p5$interval, sep=""), "%H%M"), "%H:%M")`** AM.

### 6.Impute Missing Data
``` {r Missing}
miss <- sum(is.na(df$steps))

```
There are **`r miss`** missing values in the data set for steps. The missing values are imputed using the average value for the same 5 minutes interval which are available. The data and histogram bellow is for the data set after imputing the missing values.
        
```{r Impute Missing values}
# Impute
df_new <- df[is.na(df$steps),]
df_new$steps <- 0
p4 <- with(df, aggregate(steps ~ interval, FUN = mean, na.rm = TRUE))
df_new <- transform(df_new, steps = p4$steps[match(interval, p4$interval)])
d <- df[!(is.na(df$steps)),]
df_new <-rbind(d, df_new)

# Compute Mean, Meadian
p6 <- with(df_new, aggregate(steps ~ date, FUN = mean, na.rm = TRUE))
names(p6) <- c("Date", "Mean")
p6_tmp <- with(df_new, aggregate(steps ~ date, FUN = median, na.rm = TRUE))
names(p6_tmp) <- c("Date", "Median")
p6 <- merge(p6, p6_tmp, by = "Date")

print(p6, type="HTML")

diff <- round((sum(df_new$steps, na.rm=TRUE) - sum(df$steps, na.rm=TRUE)) / sum(df_new$steps, na.rm=TRUE) * 100, 2)
```


**Observation :** There were 8 days of data that were missing intitially. Post imputing we could see the data for those 8 days and an increase in total steps by **`r diff`**


### 7.Histogram for Imputed data
``` {r Hist Imputed}
# Plot Hist
p7 <- with(df_new, aggregate(steps ~ date, FUN = sum, na.rm = TRUE))
hist(p7$steps, main="Histogram of Steps Per Day", xlab = "Total number of Steps per day (Imputed)", col = "tan")

```


### 8.Average Step Activity (Weekday Vs Weekend)
```{r Weekday Weekend Step Activity}
library(dplyr)
library(lattice)

df_new$daytype <- ifelse( weekdays(df_new$date) %in% c("Saturday", "Sunday") , "Weekend", "Weekday")
p8 <- df_new %>% select (daytype, interval, steps) %>% group_by(daytype, interval) %>% summarize ("steps" = mean(steps))
p8 <- transform(p8, daytype = factor(daytype))

xyplot (steps ~ interval | daytype, data =p8, layout = c(1,2), type = "l", xlab="Interval", ylab="Number of Steps", col="red")


```
